---
title: ""
subtitle: "Co-kriging prediction using the gstat library"
output: html_document
header-includes:
   - \usepackage{setspace}
---

#### Install packages.

```{r message=FALSE, warning=FALSE}

install.packages("raster")
install.packages("classInt")
install.packages("RColorBrewer")
install.packages("gstat")
install.packages("sf")

```


#### Load packages.

```{r message=FALSE, warning=FALSE}

library(raster)
library(classInt)
library(RColorBrewer)
library(gstat)
library("sf")
library(ggplot2)

```

#### Set working directory

Note that in R the bar of a path is /, while in Windows it is the counterbar.

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

setwd("../Solucion")

```


#### Loading the observations

Loading the shape point file which contains the observed data.

```{r message=FALSE, warning=FALSE}

dat <- read_sf("Datos_cp4/precip_seasonal.shp")

#autumn and winter average
dat$autum_winter <-(dat$autumn + dat$winter)/2

#Transform it into a sf object
dat_sf <- st_as_sf(dat)
dat <- dat_sf
plot(dat[,12:18])

```

#### Loading the Valencian Community (VC) administrative boundaries

Loading the shape polygons file which contains the border of the Valencian Community.


```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

lim_cv <- read_sf("Datos_cp4/cv.shp")
str(lim_cv)

ggplot(lim_cv) +
  geom_sf(fill = "lightblue", color = "darkblue", linewidth = 0.3) +
  theme_minimal() +
  labs(title = "Comunidad Valenciana") +
  theme(plot.title = element_text(face = "bold", size = 12),
        axis.text = element_text(size = 5)
)

```

#### Loading the raster file with the Digital Elevation Model (DEM)

Reading the raster file which contains DEM.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

ras <- raster("Datos_cp4/mde_100.tif", layer=0, values=TRUE)
str(ras)

slope <- terrain(ras, opt = "slope")
aspect <- terrain(ras, opt = "aspect")
hill <- hillShade(slope,
                  aspect,
                  angle = 45,
                  direction = 315)

par(bty = "n")
plot(hill, 
     col= grey(0:100/100),
     legend= FALSE,
     axes= FALSE,
     main= "MDT with hillshade", 
     cex.main= 1)

plot(ras,
     col= rev(terrain.colors(100, alpha = 0.5)),
     add= TRUE,
     axis.args= list(cex.axis = 0.7))

```

#**************************************************************************************************************
#### Obtaining the predicting locations
#**************************************************************************************************************

In order to obtain a continuous map of precipitation, we need to make punctual predictions in a dense grid of locations.
First of all, the spatial resolution of the raster object can be reduced.
Finally, the grid of points will be clipped by the border of CV.

1. Reducing the spatial resolution of the raster. Reduction factor=1/16.

```{r message=FALSE, warning=FALSE}

ras <- aggregate(ras, fact=c(4,4), fun=mean)
str(ras)

```

2. Converting the raster to a SpatialPoints, thus obtaing the grid of predicting points, stored in the new object 'loc'.

```{r message=FALSE, warning=FALSE}

loc <- rasterToPoints(ras, fun=NULL, spatial=TRUE)
str(loc)

```

3. Cliping the object 'loc' by the VC boundary. The resulting object 'loc' will be used as prediction locations.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

loc_sf <- st_as_sf(loc, coords = c("x", "y"), remove = FALSE)
st_crs(lim_cv) <-loc@proj4string@projargs

# Intersection loc, lim_cv
puntos_dentro <- st_intersection(loc_sf, lim_cv)

# Get the points inside the polygon
loc2 <- loc_sf[puntos_dentro, ]

loc <- as(loc2, "Spatial")
str(loc)

plot(st_geometry(lim_cv),
     col = "grey95",
     border = "black",
     main = "Sampling locations",
     cex.main = 1)

plot(loc_sf,
     add = TRUE,
     pch = 16,
     col = "lightblue",
     cex = 0.1)

plot(loc,
     add = TRUE,
     pch = 16,
     col = "darkblue",
     cex = 0.2)

legend("right",
       legend = c("All stations", "Stations within study area"),
       pch = c(1, 16),
       col = c("lightblue", "darkblue"),
       bty = "n")

```

#**************************************************************************************************************
#### Coordinate scaling
#**************************************************************************************************************
With the 'gstat' library it does not need to standardize the XY coordinates unlike the previously used library 'geoR' in the kriging model. So, we use the original units for the X and Y coordinates of both objects 'dat' and 'loc'.

#**************************************************************************************************************
#### Reference coordinate system
#**************************************************************************************************************

Set the reference coordinate system for the object 'dat', making it equal to that in the object 'loc'.

```{r message=FALSE, warning=FALSE}

#proj4string(dat) <- loc@proj4string@projargs
st_crs(dat) <-loc@proj4string@projargs
str(loc)

```
#**************************************************************************************************************
#### Initialization of the object 'gstat':
#**************************************************************************************************************

Recursive creation of the object 'g' containing the information for the co-kriging model.

First, we define it with the information of the variable 'autum_winter' and then with the information of the variable 'ALTITUD'.

The global tendency is specified by the parameter 'form' (form = spring ~ X*Y). If we do not use tendency in our model, the parameter 'form' must be set to 'form = spring ~ 1'. The parameter 'data' speficies the actual data.

```{r message=FALSE, warning=FALSE}

datos<-data.frame(dat$X,dat$Y,dat$autum_winter,dat$ALTITUD)

names(datos)<-c("X","Y","autum_winter","ALTITUD")

attach(datos)

coordinates(datos) = ~X+Y

#g <- gstat(NULL, form= spring ~ X*Y, locations= ~X+Y, data=datos)
#g <- gstat(g, form= ALTITUD ~ X*Y, locations= ~X+Y, data=datos)
g <- gstat(NULL, form= autum_winter ~ X*Y, data=datos)#, id='autum_winter')
g <- gstat(g, form= X ~ X*Y, data=datos)#, id='Xcoord')
g

```
#**************************************************************************************************************
#### Empiric and analytic variogram of the first variable
#**************************************************************************************************************

```{r message=FALSE, warning=FALSE}

plot_semivariogram <- function (variogram, xlim= NULL, ylim= NULL, model= NULL, title= NULL){
  
  plot(variogram$dist, variogram$gamma,
       pch = 20,
       xlab = "Distance",
       ylab = "Semivariance",
       main = title,
       cex.main = 1.1,
       cex.lab = 0.95,
       cex.axis = 0.95,
       xlim = if (!is.null(xlim)) xlim,
       ylim = if (!is.null(ylim)) ylim
  )

  text(x = variogram$dist,
       y = variogram$gamma ,
       labels = variogram$np,
       col= "darkblue",
       pos = 3,
       cex = 0.8
  )
  
  if (!is.null(model)) {
    line_model <- variogramLine(model, maxdist = max(variogram$dist))
     lines(line_model$dist, line_model$gamma, col = "green", lwd = 1.5)
  }
  
}

```

Empiric variogram with tendency.

```{r fig.height=4, fig.width=9, message=FALSE, warning=FALSE}

xlim= c(0, 135000)
ylim= c(0, 1500)

varig_1 <- variogram(autum_winter ~ X*Y, data=datos)

plot_semivariogram(varig_1,
                   xlim= xlim,
                   ylim= ylim,
                   title= "Semivariogram - precipitation")

```

Fit an analytical function for the variogram.


```{r message=FALSE, warning=FALSE}

get_fitted_models <- function(variogram, sill, range, nugget, names){
  
  fitted_models <- list()
  
  for (i in seq_along(names)) {
    fitted_models[[i]] <- fit.variogram(variogram, model= vgm(sill,names[i],range,nugget),fit.method= 6)
  }
  
  names(fitted_models) <- names
  
  return(fitted_models)
}

```


```{r fig.height=4*3, fig.width=9, message=FALSE, warning=FALSE}

par(mfrow=c(3,1))

sill <- 1200 
range <- 90000
nugget <- 200

names <- c("Gau", "Exp", "Sph")

models <- get_fitted_models(varig_1, sill, range, nugget, names)
models

par(mfrow=c(3,1))
for (i in seq_along(names)) {
  models[i]
  plot_semivariogram(varig_1,
                     xlim= xlim,
                     ylim= ylim,
                     model= models[[i]],
                     title= paste0("Semivariogram with - ", names[i]," - model")
  )
}

```

#**************************************************************************************************************
#### Empiric and analytic variogram of the second variable
#**************************************************************************************************************

Empiric variogram with tendency.

```{r fig.height=4, fig.width=9 message=FALSE, warning=FALSE}

xlim= c(0, 135000)
ylim= c(0, 700000)

varig_2 <- variogram(X ~ X*Y, data=datos)

plot_semivariogram(varig_2,
                   xlim= xlim,
                   ylim= ylim,
                   title= "Semivariogram - XCoord"
)

```

# Fit an analytical function for the variogram.

```{r message=FALSE, warning=FALSE}

get_fitted_models <- function(variogram, sill, range, nugget, names){
  
  fitted_models <- list()
  
  for (i in seq_along(names)) {

    fitted_models[[i]] <- fit.variogram(variogram, model= vgm(sill,names[i],range,nugget),
                                     fit.sills= TRUE, fit.ranges= TRUE, fit.method= 7,
                                     debug.level= 1, warn.if.neg= TRUE, fit.kappa= FALSE
                                     )
  }
  
  names(fitted_models) <- names
  
  return(fitted_models)
}
```


```{r fig.height=4*3, fig.width=9, message=FALSE, warning=FALSE}

sill <- 60000 
range <- 100000
nugget <- 1

models <- get_fitted_models(varig_2, sill, range, nugget, names)
models

par(mfrow=c(3,1))

for (i in seq_along(names)) {

  plot_semivariogram(varig_2,
                     xlim= xlim,
                     ylim= ylim,
                     model= models[[i]],
                     title= paste0("Semivariogram with - ", names[i]," - model")
  )
}

```

#**************************************************************************************************************
#### Cross variogram
#**************************************************************************************************************

Joint specification of both individual and cross variograms by using the same variogram model for both.

```{r fig.height=11, fig.width=9*2, message=FALSE, warning=FALSE}

sill <- 60000 
range <- 100000
nugget <- 1

model <- get_fitted_models(varig_2, sill, range, nugget, c("Gau"))
g <- gstat(g, id= "autum_winter", model= model$Gau, fill.all= T)

model$Gau
g

par(mfrow=c(2,2))

variograms <- variogram(g)

ids <- rev(unique(variogram(g)$id))

for (id in ids){
  v_sub <- variograms[variograms$id == id, ]
  plot_semivariogram(v_sub,
                     xlim= xlim,
                     ylim= c(min(v_sub$gamma), max(v_sub$gamma)),
                     title= paste0("Semivariogram - ", id)
                     )
}

```

Cross variogram

```{r message=FALSE, warning=FALSE}

varig.cross <- variogram(g)

```

Fit an analytical function for both individual and cross variograms.

```{r fig.height=11, fig.width=9*2, message=FALSE, warning=FALSE}

g <- fit.lmc(varig.cross, g, fit.method = 7, fit.ranges = FALSE, fit.lmc = TRUE, correct.diagonal = 0.95)

par(mfrow=c(2,2))

variograms <- variogram(g)
models <- g$model

ids <- rev(unique(variogram(g)$id))

for (id in ids){
  v_sub <- variograms[variograms$id == id, ]
  current_model <- models[[id]] 
  plot_semivariogram(v_sub,
                     xlim= xlim,
                     ylim= c(min(v_sub$gamma), max(v_sub$gamma)),
                     model= current_model,
                     title= paste0("Semivariogram - ", id)
                     )
}

```


#**************************************************************************************************************
#### Co-krigin prediction
#**************************************************************************************************************
we need to have the same names for variables in both objects 'dat' and 'loc'.

```{r message=FALSE, warning=FALSE}

#loc@data
loc@data$ALTITUD <- loc@data$mde_100 
loc@data$X <- loc@coords[,1]
loc@data$Y <- loc@coords[,2]
loc1 <- data.frame(loc@data$mde_100,loc@coords[,1],loc@coords[,2])
names(loc1)<-c("ALTITUD","X","Y")
coordinates(loc1) = ~X+Y

```

Co-kriging function

```{r message=FALSE, warning=FALSE}

#kc <- predict(g, newdata=loc@data)
kc <- predict(g, newdata=loc1)
str(kc)

loc@data$predict <- kc$var1.pred   #predictions
loc@data$predict.sd <- sqrt(kc$var1.var)   #predicting errors

```

Histograms

```{r fig.height=6, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}

par(mfrow=c(2,2))

hist(kc$var1.pred, breaks=30, main="predictions", xlab= "precipitation (mm)") 
hist(sqrt(kc$var1.var), breaks=30, main="prediction errors", xlab= "error (mm)")
hist(kc$cov.var1.var2, breaks=30, main="co-semivariances", xlab= "co-semivariances")

```
#**************************************************************************************************************
#### Cross-validation
#**************************************************************************************************************
Cross-validation
 
```{r  message=FALSE, warning=FALSE}

kc.cv <- gstat.cv(g, remove.all = FALSE, all.residuals = FALSE, maxdist=120000, nmax=150, nfold= 30)

str(kc.cv)

```

Histograms

```{r fig.height=6, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}

summary(kc.cv$residual)

par(mfrow=c(2,2))

# Histogram of the cross-validation

hist(kc.cv$observed, breaks=30, main="Observations", xlab="precipitation (mm)")
hist(kc.cv$var1.pred, breaks=30, main="CV-predictions", xlab="precipitation (mm)") 
hist(kc.cv$residual, breaks=30, main="CV-residuals", xlab="residual (mm)") 
hist(sqrt(kc.cv$var1.var), breaks=30, main="CV-prediction errors", xlab="error (mm)")

# Diagnosis plots

hist(kc.cv$residual, breaks=30, main="CV-residuals", xlab= "residual (mm)") 
hist(sqrt(kc.cv$var1.var), breaks=30, main="CV-prediction errors", xlab = "error (mm)") 

plot(kc.cv$observed, kc.cv$var1.pred,
     main="CV-predictions vs observed",
     xlab = "precipitation (mm)",
     ylab = "predicted precipitation (mm)",       
     cex.main = 1.1,
     cex.lab = 0.95,
     cex.axis = 0.95) 
abline(a=0, b=1, col="blue", lty=1, lwd=1) 

plot(kc.cv$observed, kc.cv$residual,
     main="CV-residuals vs observed", 
     xlab = "precipitation (mm)",
     ylab = "residuals (mm)",
     cex.main = 1.1,
     cex.lab = 0.95,
     cex.axis = 0.95)

abline(h=0, col="blue", lty=2)

```

#**************************************************************************************************************
#### Map of co-kriging prediction:
#**************************************************************************************************************

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}

par(mfrow=c(1,2))

######## Map 1 - Predictions########

# Predictions
loc2 <- loc
loc2$predict <- SpatialPixelsDataFrame(loc@coords, data=as.data.frame(loc@data$predict))
q5 <- quantile(loc2$predict@data[,1], c(0,0.10,0.25,0.40,0.60,0.75,0.90,1))

# Predicted precipitation
image(loc2$predict,
      col= brewer.pal(7,"Blues"),
      asp= 1,
      breaks= q5,
      frame.plot= FALSE,
      xlim= c(600000, 800000),
      main= "Predictions",
      cex.main= 1)

# Hillshade
cols <- adjustcolor(grey(0:100/100), alpha.f = 0.4)
image(hill, col= cols, asp= 1, add= TRUE)

# Legend
br5 <- cut(loc2$predict@data[,1], breaks=q5, include.lowest=TRUE)
legend("bottomleft",
       legend= rev(levels(br5)),
       fill= rev(brewer.pal(7,"Blues")),
       title= "Predictions (mm)",
       cex= 0.7)

# Observed data
q6 <- quantile(datos$autum_winter, c(0,0.10,0.25,0.40,0.60,0.75,0.90,1))
br6 <- cut(datos$autum_winter, breaks=q6, include.lowest=TRUE)

points(datos$X, datos$Y,
       col= brewer.pal(7,"Spectral")[7:1][br6],
       asp= 1,
       pch= 16,
       cex= 0.5)

legend("bottomright",
       legend= levels(br6)[7:1],
       fill= brewer.pal(7,"Spectral")[1:7],
       title= "Observations (mm)",
       cex= 0.7)

####### Map 2 - Predicting errors########

loc2$predict.sd <- SpatialPixelsDataFrame(loc@coords, data=as.data.frame(loc@data$predict.sd))
q7 <- quantile(loc2$predict.sd@data[,1], c(0,0.10,0.25,0.40,0.60,0.75,0.90,1))

image(loc2$predict.sd,
      col= brewer.pal(7, "BuPu"),
      asp= 1,
      breaks= q7,
      frame.plot= FALSE,
      xlim= c(600000, 800000),
      main= "Prediction errors",
      cex.main= 1)

# Hillshade
image(hill, col= cols, asp= 1, add= TRUE)

# Legend
br7 <- cut(loc2$predict.sd@data[,1], breaks=q7, include.lowest=TRUE)
legend("bottomright", 
       legend= rev(paste(levels(br7),c(" - q0.10"," - q0.25"," - q0.40"," - q0.60"," - q0.75"," - q0.90"," - q1.0"))),
       fill= rev(brewer.pal(7, "BuPu")), 
       title= "prediction errors",
       cex= 0.7)

# Observed points
points(datos$X, datos$Y,
       col= brewer.pal(7,"Spectral")[7:1][br6],
       asp= 1,
       pch= 16,
       cex= 0.5)

```
