---
title: ""
subtitle: "Kriging prediction using the ‘geoR’ library"
output: html_document
header-includes:
   - \usepackage{setspace}
---

#### Install packages.

```{r message=FALSE, warning=FALSE}

install.packages("raster")
install.packages("classInt")
install.packages("RColorBrewer")
install.packages("geoR")
install.packages("sf")

```


#### Load packages.

```{r message=FALSE, warning=FALSE}

library(raster)
library(classInt)
library(RColorBrewer)
library(geoR)
library("sf")
library(ggplot2)

```

#### Set working directory

Note that in R the bar of a path is /, while in Windows it is the counterbar.

```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

setwd("../Solucion")

```


#### Loading the observations

Loading the shape point file which contains the observed data.

```{r message=FALSE, warning=FALSE}

dat <- read_sf("Datos_cp4/precip_seasonal.shp")

#autumn and winter average
dat$autum_winter <-(dat$autumn + dat$winter)/2

#Transform it into a sf object
dat_sf <- st_as_sf(dat)
dat <- dat_sf

```

#### Loading the Valencian Community (VC) administrative boundaries

Loading the shape polygons file which contains the border of the Valencian Community.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

lim_cv <- read_sf("Datos_cp4/cv.shp")
str(lim_cv)
ggplot(lim_cv) +
  geom_sf(fill = "lightblue", color = "darkblue", linewidth = 0.3) +
  theme_minimal() +
  labs(title = "Comunidad Valenciana") +
  theme(plot.title = element_text(face = "bold", size = 12),
        axis.text = element_text(size = 5)
  )

```




#**************************************************************************************************************
#### Loading the raster file with the Digital Elevation Model (DEM)
#**************************************************************************************************************

Reading the raster file which contains DEM.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

ras <- raster("Datos_cp4/mde_100.tif", layer=0, values=TRUE)
str(ras)

slope <- terrain(ras, opt = "slope")
aspect <- terrain(ras, opt = "aspect")
hill <- hillShade(slope,
                  aspect,
                  angle = 45,
                  direction = 315)

par(bty = "n")
plot(hill, 
     col= grey(0:100/100),
     legend= FALSE,
     axes= FALSE,
     main= "MDT with hillshade", 
     cex.main= 1)

plot(ras,
     col= rev(terrain.colors(100, alpha = 0.5)),
     add= TRUE,
     axis.args= list(cex.axis = 0.7))

```

#**************************************************************************************************************
#### Obtaining the predicting locations
#**************************************************************************************************************
In order to obtain a continuous map of precipitation, we need to make punctual predictions in a dense grid of locations.
First of all, the spatial resolution of the raster object can be reduced. 
Finally, the grid of points will be clipped by the border of CV.

1. Reducing the spatial resolution of the raster. Reduction factor=1/16.

```{r message=FALSE, warning=FALSE}

ras <- aggregate(ras, fact=c(4,4), fun=mean) # Resampling
str(ras)

```

2. Converting the raster to a SpatialPoints, thus obtaing the grid of predicting points, stored in the new object 'loc'.

```{r message=FALSE, warning=FALSE}

loc <- rasterToPoints(ras, fun=NULL, spatial=TRUE)
str(loc)

```

3. Cliping the object 'loc' by the VC boundary. The resulting object 'loc' will be used as prediction locations.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

loc_sf <- st_as_sf(loc, coords = c("x", "y"), remove = FALSE)
st_crs(lim_cv) <-loc@proj4string@projargs

# Intersection loc, lim_cv
puntos_dentro <- st_intersection(loc_sf, lim_cv)

# Get the points inside the polygon
loc2 <- loc_sf[puntos_dentro, ]

loc <- as(loc2, "Spatial")
str(loc)

plot(st_geometry(lim_cv),
     col = "grey95",
     border = "black",
     main = "Sampling locations",
     cex.main = 1)

plot(loc_sf,
     add = TRUE,
     pch = 16,
     col = "lightblue",
     cex = 0.1)

plot(loc,
     add = TRUE,
     pch = 16,
     col = "darkblue",
     cex = 0.2)

legend("right",
       legend = c("All stations", "Stations within study area"),
       pch = c(1, 16),
       col = c("lightblue", "darkblue"),
       bty = "n")

```

#**************************************************************************************************************
#### Scale and center the coordinates (Normalization)
#**************************************************************************************************************

Due to the fact that when using the library 'geoR' to peformm kriging, an numerical inestability error is obtained when data have UTM spatial coordinates, we should first scale the spatial coordinates variables.

```{r message=FALSE, warning=FALSE}

media <- colMeans(loc@coords[,]) # Centroid of the object 'loc'
desv <- sd(loc@coords[,]) # Common standard deviation of X and Y coordinates.

```

Scaling the observations

```{r message=FALSE, warning=FALSE}

#dat@coords <- scale(dat@coords, center= media, scale= c(desv,desv))
#str(dat)
dat$X1 = (dat$X-media[1])/desv
dat$Y1 = (dat$Y-media[2])/desv

```

Scaling the predicting locations

```{r message=FALSE, warning=FALSE}

loc@coords <- scale(loc@coords,
                    center= media,
                    scale= c(desv,desv))
str(loc)

```

#**************************************************************************************************************
#### Isotropic (omnidirectional) empiric variogram.
#**************************************************************************************************************

```{r message=FALSE, warning=FALSE}

add_models <- function(models) {
  cols <- c("red", "darkblue", "cyan")
  for (i in seq_along(models)) {
    lines(models[[i]], lwd = 1, col = cols[i])
  }
  legend("bottomright", title = "Model", legend = c("exponential", "gaussian", "spherical"),
         col = cols, lty = 1, lwd = 1.5, cex = 0.7)
}

plot_semivariogram <- function(svar1, xlabel1, svar2, xlabel2, desv=NULL, models=NULL) {
  
  par(mfrow=c(1,2))
  
  #First graph
  plot(svar1, pch=20, 
       xlab=xlabel1,
       ylab = "Semivariance",
       main = "Semivariogram",
       cex.main = 1.1,
       cex.lab = 0.95,
       cex.axis = 0.95,
       xaxt = if (is.null(desv)) "s" else "n"
  )
  
  if (!is.null(desv)) {  
    axis(1, at= c(0,50000,150000,250000,350000)/desv,
         labels= c(0,50000,150000,250000,350000),
         tick= TRUE,
         lty= 1,
         las=1,
         cex.axis=1,
         font=1,
         col=grey(0.5),
         col.ticks=grey(0.3))
  }
    
  # Add models
  if (!is.null(models)) {
    add_models(models)
  }
  
  #Second graph
  plot(svar2, pch=20, 
       xlab=xlabel2,
       ylab = "Semivariance",
       main = "Semivariogram",
       cex.main = 1.1,
       cex.lab = 0.95,
       cex.axis = 0.95
  )
  ## Add model
  if (!is.null(models)) {
    add_models(models)
  }
  
  par(mfrow=c(1,1))

}

```


Semivariogram without tendency

```{r fig.height=4, fig.width=9, message=FALSE, warning=FALSE}

svar1 <- variog(coords=data.frame(dat$X1,dat$Y1),
                data=dat$autum_winter,
                uvec=seq(0,0.19,0.005),
                trend= "cte")  
str(svar1)

svar1_undo <- svar1
svar1_undo$u <- svar1$u*desv
 
plot_semivariogram(svar1 =  svar1_undo,
                   xlabel1 = "original distance",
                   svar2 = svar1,
                   xlabel2 = "scaled distance"
)

```

Semivariogram with tendency

```{r fig.height=4, fig.width=9, message=FALSE, warning=FALSE}

svar2 <- variog(coords=data.frame(dat$X1,dat$Y1),
                data=dat$autum_winter,
                uvec=seq(0,0.19,0.005),
                trend= ~dat$X1*dat$Y1*dat$ALTITUD)
str(svar2)

plot_semivariogram(svar1 = svar2,
                   xlabel1 = "original distance",
                   svar2 = svar2,
                   xlabel2 = "scaled distance",
                   desv = desv
)
```
### Histogram of distances

```{r fig.height=4, fig.width=9, message=FALSE, warning=FALSE}

par(mfrow=c(1,2))

########## Plot with the original distances

distancias_undo <- as.matrix(dist(data.frame(dat$X1,dat$Y1)*desv))[upper.tri(as.matrix(dist(data.frame(dat$X1,dat$Y1))), diag=FALSE)]

hist(distancias_undo[distancias_undo<=0.19*desv],
     breaks=seq(0,0.19*desv,0.005*desv),
     main="", xlab="original distance"
)

########## Plot with the scaled distances
distancias <- as.matrix(dist(data.frame(dat$X1,dat$Y1)))[upper.tri(as.matrix(dist(data.frame(dat$X1,dat$Y1))), diag=FALSE)]
                        
hist(distancias[distancias<=0.19],
     breaks=seq(0,0.19,0.005),
     main="",
     xlab="scaled distance"
)

```

#**************************************************************************************************************
#***********************************************VARIOGRAM MODEL************************************************
#***********************************Adjust of different variogram functions************************************
#**************************************************************************************************************


```{r message=FALSE, warning=FALSE}

fit_variogram <- function(vario, sill, range, nugget, covmodel) {
  variofit(
    vario = vario,
    ini = c(sill, range),
    cov.model = covmodel,
    fix.nugget = FALSE,
    nugget = nugget,
    weights = "cressie",
    minimisation.function = "optim",
    max.dist = 0.2,
    fix.kappa = FALSE
  )
}

```



```{r fig.height=4, fig.width=9, message=FALSE, warning=FALSE}

sill <- 1180 
range <- 0.05  # Notice it is in the scaled range
nugget <- 200

svar.exp <- fit_variogram(svar2, sill, range, nugget, "exponential")
svar.gau <- fit_variogram(svar2, sill, range, nugget, "gaussian")
svar.sph <- fit_variogram(svar2, sill, range, nugget, "spherical")

plot_semivariogram(svar1 = svar2,
                   xlabel1 = "original distance",
                   svar2 = svar2,
                   xlabel2 = "scaled distance",
                   desv = desv,
                   models = list(svar.exp, svar.gau, svar.sph)
)

```

#**************************************************************************************************************
#### Kriging adjustment
#**************************************************************************************************************
Adjustment of Kriging model 

```{r message=FALSE, warning=FALSE}

uk <- krige.conv(data=dat$autum_winter, coords=data.frame(dat$X1,dat$Y1), 
                 locations= loc@coords, 
                 krige= krige.control(type.krige = "ok", trend.d= trend.spatial( ~dat$X1*dat$Y1*dat$ALTITUD ),
                                      trend.l= trend.spatial( ~loc@coords[,1]*loc@coords[,2]*loc@data$mde_100 ),
                                      obj.model= svar.exp))  
str(uk)

```

Adding kriging outcomes to the object 'loc'

```{r message=FALSE, warning=FALSE}

loc@data$predict <- uk$predict  # predictions
loc@data$predict.sd <- sqrt(uk$krige.var)  # prediction error

```

#**************************************************************************************************************
#### Cross-validation (CV) model assesment
#**************************************************************************************************************

```{r echo=True, message=FALSE, warning=FALSE, results=FALSE, paged.print=FALSE}

for(i in 1:212){
  
  uk_vc <- krige.conv(data= dat$autum_winter[-i],
                      coords= data.frame(dat$X1,dat$Y1)[-i,],
                      locations= data.frame(dat$X1,dat$Y1)[i,],
                      borders= NULL,
                      krige= krige.control(type.krige= "ok",
                                           trend.d= trend.spatial( ~dat$X1[-i]*dat$Y1[-i]*dat$ALTITUD[-i]),
                                           trend.l= trend.spatial( ~dat$X1[i]*dat$Y1[i]*dat$ALTITUD[i]),
                                           obj.model= svar.exp))
  
  dat$predict.cv[i] <- uk_vc$predict  
  dat$predict.cv.sd[i] <- sqrt(uk_vc$krige.var)  
}	

```


Computing Cross-validation (CV) residuals

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

dat$res.cv <- dat$autum_winter[] - dat$predict.cv

str(dat)

```

Diagnosis plots


```{r message=FALSE, warning=FALSE}

plot_diagnostics <- function(obs, model) {

  par(mfrow = c(2, 2))

  # Histogram fitted residuals
  hist(model$residuals,
       breaks = 20,
       main = "Residuals",
       xlab = "Residual")

  # Histogram observations
  hist(obs,
       breaks = 20,
       main = "Observations",
       xlab = "Precipitation (mm)")

  # Plot observations vs fitted values
  plot(obs, model$fitted.values,
       asp = 1,
       xlab = "Precipitation (mm)",
       ylab = "Predicted precipitation (mm)",
       main = "Observation vs Prediction",
       cex.main = 1.1,
       cex.lab = 0.95,
       cex.axis = 0.95)

  abline(a = 0, b = 1, col = "blue")
  
  # Plot observations vs residuals
  plot(obs, model$residuals,
       asp = 1,
       xlab = "Precipitation (mm)",
       ylab = "Residuals",
       main = "Observation vs Residual",
       cex.main = 1.1,
       cex.lab = 0.95,
       cex.axis = 0.95)

  abline(h = 0, col = "blue", lty = 2)

  par(mfrow = c(1, 1))
}
```






```{r fig.height=6, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}

#Summary of CV-residuals
summary(dat$res.cv)

# Histogram of the cross-validation
par(mfrow=c(2,2))
hist(dat$autum_winter, breaks=30, main="Observations", xlab="precipitation (mm)")
hist(dat$predict.cv, breaks=30, main="CV-predictions (over training data)", xlab="precipitation (mm)") 
hist(dat$res.cv, breaks=30, main="CV-residuals", xlab="residual (mm)")  
hist(dat$predict.cv.sd, breaks=30, main="CV-prediction errors", xlab="error (mm)")   

# Diagnosis plots
par(mfrow=c(2,2))
hist(dat$res.cv, breaks=30, main="CV-residuals", xlab= "residual (mm)") 
hist(dat$predict.cv.sd, breaks=30, main="CV-prediction errors", xlab = "error (mm)") 

plot(dat$autum_winter, dat$predict.cv,
     main="CV-predictions vs observed",
     xlab = "precipitation (mm)",
     ylab = "predicted precipitation (mm)",       
     cex.main = 1.1,
     cex.lab = 0.95,
     cex.axis = 0.95) 
abline(a=0, b=1, col="blue", lty=1, lwd=1) 

plot(dat$autum_winter, dat$res.cv,
     main="CV-residuals vs observed", 
     xlab = "precipitation (mm)",
     ylab = "residuals (mm)",
     cex.main = 1.1,
     cex.lab = 0.95,
     cex.axis = 0.95)

abline(h=0, col="blue", lty=2) 

```

Identifying outliers

```{r message=FALSE, warning=FALSE}

dat[ dat$res.cv < -40 | dat$res.cv > 40 , c(1:2,12:19)]

```


#**************************************************************************************************************
#### Undo the scaling of the spatial coordinates
#**************************************************************************************************************

```{r message=FALSE, warning=FALSE}

loc@coords <- scale(loc@coords, center=FALSE, scale= 1/c(desv,desv))
loc@coords <- scale(loc@coords, center= -1*media, scale=FALSE)

dat$X2 = (dat$X1*desv)+media[1]
dat$Y2 = (dat$Y1*desv)+media[2]

#dat@coords <- scale(dat@coords, center=FALSE, scale= 1/c(desv,desv))
#dat@coords <- scale(dat@coords, center= -1*media, scale=FALSE)

```

#**************************************************************************************************************
#### Map of the kriging predictions
#**************************************************************************************************************
```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}

par(mfrow=c(1,2))

######## Map 1 - Predictions########

loc1_predict <- SpatialPixelsDataFrame(loc@coords, data=as.data.frame(loc@data$predict))
q2 <- quantile(loc1_predict@data[,1], c(0,0.10,0.25,0.40,0.60,0.75,0.90,1))

# Predicted precipitation
image(loc1_predict,
      col= brewer.pal(7,"Blues"),
      asp= 1,
      breaks= q2,
      frame.plot= FALSE,
      xlim= c(600000, 800000),
      main= "Predictions",
      cex.main= 1)

# Hillshade
cols <- adjustcolor(grey(0:100/100), alpha.f = 0.4)
image(hill, col= cols, asp= 1, add= TRUE)

# Legend
br2 <- cut(loc1_predict@data[,1], breaks=q2, include.lowest=TRUE)
legend("bottomleft",
       legend= rev(levels(br2)),
       fill= rev(brewer.pal(7,"Blues")),
       title= "Predictions (mm)",
       cex= 0.7)


# Observed data.
q3 <- quantile(dat$autum_winter[], c(0,0.10,0.25,0.40,0.60,0.75,0.90,1))
br3 <- cut(dat$autum_winter[], breaks=q3, include.lowest=TRUE)

points(dat$X2, dat$Y2,
       col= brewer.pal(7,"Spectral")[7:1][br3],
       asp= 1,
       pch= 16,
       cex= 0.5)

legend("bottomright",
       legend= levels(br3)[7:1],
       fill= brewer.pal(7,"Spectral")[1:7],
       title= "Observations (mm)",
       cex= 0.7)


####### Map 2 - Predicting errors########

loc1_predict.sd <- SpatialPixelsDataFrame(loc@coords, data=as.data.frame(loc@data$predict.sd))
q4 <- quantile(loc1_predict.sd@data[,1], c(0,0.10,0.25,0.40,0.60,0.75,0.90,1))

image(loc1_predict.sd,
      col= brewer.pal(7, "BuPu"),
      asp= 1,
      breaks= q4,
      frame.plot= FALSE,
      xlim= c(600000, 800000),
      main= "Prediction errors",
      cex.main= 1)

# Hillshade
image(hill, col= cols, asp= 1, add= TRUE)

# Legend
br4 <- cut(loc1_predict.sd@data[,1], breaks=q4, include.lowest=TRUE)
legend("bottomright", 
       legend= rev(paste(levels(br4),c(" - q0.10"," - q0.25"," - q0.40"," - q0.60"," - q0.75"," - q0.90"," - q1.0"))),
       fill= rev(brewer.pal(7, "BuPu")), 
       title= "prediction errors",
       cex= 0.7)

# Observed data.
points(dat$X2, dat$Y2,
       col= brewer.pal(7,"Spectral")[7:1][br3],
       asp= 1,
       pch= 16,
       cex= 0.5)

```


