---
title: ""
subtitle: "Descriptive analysis and linear regressions"
output: html_document
header-includes:
   - \usepackage{setspace}
---

#### Install packages.

```{r message=FALSE, warning=FALSE}

install.packages("raster")
install.packages("classInt")
install.packages("RColorBrewer")
install.packages("sf")
install.packages("ggplot2")

```


#### Load packages.

```{r message=FALSE, warning=FALSE}

library(raster)
library(classInt)
library(RColorBrewer)
library("sf")
library(ggplot2)

```

#### Set working directory
```{r message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE}

setwd("../Solucion")

```
#### Loading and ploting the observations

Reading the shape point file which contains the data and converting it to a spatial-points object of R.

```{r message=FALSE, warning=FALSE}

dat <- read_sf("Datos_cp4/precip_seasonal.shp") # (from shp to object of R)

#autumn and winter average
dat$autum_winter <-(dat$autumn + dat$winter)/2

#Transform it into a sf object
dat_sf <- st_as_sf(dat)
dat <- dat_sf

```

A summary of the structure and content of the created object

```{r message=FALSE, warning=FALSE}

str(dat)

```


```{r message=FALSE, warning=FALSE}

dev.new()
plot(dat[,12:18])

```

#### Loading and ploting the observations the Valencian Community (VC) administrative boundaries
Reading the shape polygons file which contains the border of the Valencian Community and converting it to a spatial-polygons object of R.

Visualizing a summary of the structure and content of the object.

Plot the border map.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

#lim_cv <- readShapePoly("Datos_cp4/cv.shp")
lim_cv <- read_sf("Datos_cp4/cv.shp")
str(lim_cv)
ggplot(lim_cv) +
  geom_sf(fill = "lightblue", color = "darkblue", linewidth = 0.3) +
  theme_minimal() +
  labs(title = "Comunidad Valenciana") +
  theme(plot.title = element_text(face = "bold", size = 12),
        axis.text = element_text(size = 5)
)

```

#### Loading and ploting the the raster file with the Digital Elevation Model (DEM)

Reading the raster file which contains DEM and converting it to a raster object of R.

Visualizing a summary of the structure and content of the object.

Plot the DEM map.

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

ras <- raster("Datos_cp4/mde_100.tif", layer=0, values=TRUE)
str(ras)

slope <- terrain(ras, opt = "slope")
aspect <- terrain(ras, opt = "aspect")
hill <- hillShade(slope, aspect, angle = 45, direction = 315)

par(bty = "n")
plot(hill, col = grey(0:100/100), legend = FALSE, axes = FALSE, main = "MDT with hillshade",  cex.main = 1)
plot(ras, col = rev(terrain.colors(100, alpha = 0.5)), add = TRUE, axis.args = list(cex.axis = 0.7))

```
#### Plot of the observations and the DEM

```{r fig.height=7, fig.width=7, message=FALSE, warning=FALSE}

par(cex.axis = 0.7)

# ploting the DEM
q1 <- classIntervals(getValues(ras)[getValues(ras)>0.05], n = 15, style = "quantile")
q1$brks <- round(c(0,q1$brks),1)
col <- c("#deebf7", gray.colors(16, start = 0.0, end = 1, alpha = 1)[-1])

image(hill, col = grey(0:100/100), asp = 1, frame.plot = FALSE, xlim = c(600000, 1000000),
      main = "Average Fall-Winter Precipitation", cex.main = 1)
image(ras, col = col, breaks = q1$brks, asp = 1, frame.plot = FALSE, xlim = c(600000, 1000000), add = TRUE)

# adding the DEM's legend
br <- cut(getValues(ras), breaks=q1$brks, include.lowest=TRUE)
legend(x=850000, y=4370000, legend=levels(br)[15:1], fill=col[15:1], ncol=1, title= "MDT", cex=0.7, box.col = "grey")

# adding the borders of VC.
lines(lim_cv, col="darkblue", lwd=0.3)

# adding the observations
q <- quantile(dat$autum_winter, c(0,0.10,0.25,0.40,0.60,0.75,0.90,1))
br1 <- cut(dat$autum_winter, breaks=q, include.lowest=TRUE)
points(dat$X, dat$Y, col=brewer.pal(7,"Spectral")[7:1][br1], asp=1, pch=16) 

# adding a legend for the observations
legend(x=850000, y=4500000, legend=levels(br)[7:1], fill=brewer.pal(7,"Spectral")[1:7], title= "Autum_winter precipitation", cex=0.7, box.col = "grey")

```
#**************************************************************************************************************
#******************************Frecuency histograms of the precipitation variables*****************************
#**************************************************************************************************************

Frecuency histograms and standard deviation of the precipitation variables: 'annual', 'spring', 'summer', 'autumn', 'winter' and 'autum_winter'.

```{r message=FALSE, warning=FALSE}

par(mfrow=c(2,3))

hist(dat$annual, breaks=20, main="annual", xlab="Precipitation (mm)"); print("sd annual"); sd(dat$annual) 
hist(dat$spring, breaks=20, main="spring", xlab="Precipitation (mm)"); print("sd spring"); sd(dat$spring)
hist(dat$summer, breaks=20, main="summer", xlab="Precipitation (mm)"); print("sd summer"); sd(dat$summer)
hist(dat$autumn, breaks=20, main="autumn", xlab="Precipitation (mm)"); print("sd autumn"); sd(dat$autumn)
hist(dat$winter, breaks=20, main="winter", xlab="Precipitation (mm)"); print("sd winter"); sd(dat$winter)
hist(dat$autum_winter, breaks=20, main="autum_winter", xlab="Precipitation (mm)"); print("sd autum_winter"); sd(dat$autum_winter)

```

#### Frecuency histograms of co-variables.

Frecuency histograms of the variables: 'ALTITUD', 'Xutm', 'Yutm'.

```{r message=FALSE, warning=FALSE}

par(mfrow=c(2,2))

hist(dat$ALTITUD, breaks=20, main="ALTITUDE", xlab="Altitude(m)") 
hist(sqrt(dat$ALTITUD), breaks=20, main="sqrt(ALTITUDE)", xlab="Square root of the altitude") 
hist(dat$X, breaks=20, main="coord X", xlab="X coordinate (m)")
hist(dat$Y, breaks=20, main="coord Y", xlab="Y coordinate (m)")

```
#**************************************************************************************************************
#****************************************Relationship between variables****************************************
#**************************************************************************************************************


```{r message=FALSE, warning=FALSE}

par(mfrow = c(1, 3))

plot(dat$autum_winter, dat$ALTITUD,
     xlab = "autum-winter precipitation (mm)",
     ylab = "Altitude (m)",
     main = "Precipitation vs altitude",
     cex.main = 1.1, cex.lab = 0.95, cex.axis = 0.95,
)
correl = round(cor(dat$autum_winter, dat$ALTITUD),3)
mtext(paste0("ρ = ", correl), side = 3, line = 0.3, cex = 0.6)

plot(dat$autum_winter, dat$X,
     xlab = "autum-winter precipitation (mm)",
     ylab = "X coordinate (m)",
     main = "Precipitation vs X coordinate",
     cex.main = 1.1, cex.lab = 0.95, cex.axis = 0.95,
)
correl = round(cor(dat$autum_winter, dat$X),3)
mtext(paste0("ρ = ", correl), side = 3, line = 0.3, cex = 0.6)

plot(dat$autum_winter, dat$Y,
     xlab = "autum-winter precipitation (mm)",
     ylab = "y coordinate (m)",
     main = "Precipitation vs Y coordinate",
     cex.main = 1.1, cex.lab = 0.95, cex.axis = 0.95,
)
correl = round(cor(dat$autum_winter, dat$Y),3)
mtext(paste0("ρ = ", correl), side = 3, line = 0.3, cex = 0.6)

par(mfrow = c(1, 1))

```

#**************************************************************************************************************
#**********************************************LINEAR REGRESION MODELS******************************
#**************************************************************************************************************


```{r message=FALSE, warning=FALSE}

plot_diagnostics <- function(obs, model) {

  par(mfrow = c(2, 2))

  # Histogram fitted residuals
  hist(model$residuals,
       breaks = 20,
       main = "Residuals",
       xlab = "Residual")

  # Histogram observations
  hist(obs,
       breaks = 20,
       main = "Observations",
       xlab = "Precipitation (mm)")

  # Plot observations vs fitted values
  plot(obs, model$fitted.values,
       asp = 1,
       xlab = "Precipitation (mm)",
       ylab = "Predicted precipitation (mm)",
       main = "Observation vs Prediction",
       cex.main = 1.1,
       cex.lab = 0.95,
       cex.axis = 0.95)

  abline(a = 0, b = 1, col = "blue")
  
  # Plot observations vs residuals
  plot(obs, model$residuals,
       asp = 1,
       xlab = "Precipitation (mm)",
       ylab = "Residuals",
       main = "Observation vs Residual",
       cex.main = 1.1,
       cex.lab = 0.95,
       cex.axis = 0.95)

  abline(h = 0, col = "blue", lty = 2)

  par(mfrow = c(1, 1))
}

```

#### *Model 1* - Simple lineal with Altitude:

$$
\begin{align}
& X^{autum-winter}_i \sim \mathcal{N}(\mu_i,\sigma^2),  \hspace{15mm} i=1,\cdots,n \\
\\
&\mu_i =  \beta_0 + \beta_1 X^{altitud}_i
\end{align}
$$
#******************************
#*******Model adjustment*******
#******************************

```{r message=FALSE, warning=FALSE}

aj_1 <- lm( autum_winter ~ ALTITUD, data=dat)
summary(aj_1)

```
#******************************
#*******Diagnosis plots********
#******************************

```{r fig.height=6, fig.width=7, message=FALSE, warning=FALSE}

plot_diagnostics(dat$autum_winter, aj_1)

```

#### *Model 2* - Linear Multiple with X, Y, and the interaction between X and Y:

$$
\begin{align}
& X^{autum-winter}_i \sim \mathcal{N}(\mu_i,\sigma^2),  \hspace{15mm} i=1,\cdots,n \\
\\
&\mu_i =  \beta_0 + \beta_1 \cdot X^{Xutm}_i + \beta_2 \cdot X^{Yutm}_i + \beta_3 \cdot X^{Xutm}_i \cdot X^{Yutm}_i.
\end{align}
$$
#******************************
#*******Model adjustment*******
#******************************
```{r message=FALSE, warning=FALSE}

aj_2 <- lm( autum_winter ~ dat$X*dat$Y, data=dat)
summary(aj_2)

```

#******************************
#*******Diagnosis plots********
#******************************

```{r fig.height=6, fig.width=7, message=FALSE, warning=FALSE}

plot_diagnostics(dat$autum_winter, aj_2)

```


#### *Model 3* - Multiple Linear with X, Y, ALTITUDE, and the interaction between X and Y:

$$
\begin{align}
& X^{autum-winter}_i \sim \mathcal{N}(\mu_i,\sigma^2),  \hspace{15mm} i=1,\cdots,n \\
\\
&\mu_i =  \beta_0 + \beta_1 \cdot X^{altitud}_i + \beta_2 \cdot X^{Xutm}_i + \beta_3 \cdot X^{Yutm}_i + \beta_4 \cdot X^{Xutm}_i \cdot X^{Yutm}_i.
\end{align}
$$
#******************************
#*******Model adjustment*******
#******************************
```{r message=FALSE, warning=FALSE}

aj_3 <- lm( autum_winter ~ ALTITUD + dat$X*dat$Y , data=dat )
summary(aj_3)

```

#******************************
#*******Diagnosis plots********
#******************************

```{r fig.height=6, fig.width=7, message=FALSE, warning=FALSE}

plot_diagnostics(dat$autum_winter, aj_3)

```


#### *Model 4*: multiple linear regression with X, Y, ALTITUDE, and the interaction between the three variables:

$$
\begin{align}
& X^{autum-winter}_i \sim \mathcal{N}(\mu_i,\sigma^2),  \hspace{15mm} i=1,\cdots,n \\
\\
&\mu_i =  \beta_0 + \beta_1 \cdot X^{altitud}_i + \beta_2 \cdot X^{Xutm}_i + \beta_3 \cdot X^{Yutm}_i + \beta_4 \cdot X^{altitud}_i \cdot X^{Xutm}_i \\
& + \beta_5 \cdot X^{altitud}_i \cdot X^{Yutm}_i + \beta_6 \cdot X^{Xutm}_i \cdot X^{Yutm}_i + \beta_7 \cdot X^{altitud}_i \cdot X^{Xutm}_i \cdot X^{Yutm}_i .
\end{align}
$$

#******************************
#*******Model adjustment*******
#******************************
```{r message=FALSE, warning=FALSE}

aj_4 <- lm( autum_winter ~ ALTITUD*dat$X*dat$Y , data=dat )
summary(aj_4)

```
#******************************
#*******Diagnosis plots********
#******************************

```{r fig.height=6, fig.width=7, message=FALSE, warning=FALSE}

plot_diagnostics(dat$autum_winter, aj_4)

```


